Speakers of diverse languages structure their utterances for efficient communication

### 2019-09-13 Paper/paper.Rmd
references and r^2s

```{r setup, include=FALSE}
library(tidyverse)
library(ggdendro)
library(here)
library(feather)
library(ggdendro)
library(gridExtra)

@@

See in the figure below a qualitative comparison of the information trajectory they found and the information trajectory we found. It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as @yu2016distribution did. The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.

@@

For Switchboard we find the same information trajectory as we did for the BNC, and we find the same distribution of information regardless of utterance length. The longer the utterance, the more pronounced the information distribution shape becomes. A qualitative comparison of the two trajectories is below.

```{r, echo=F, fig.height=2, fig.cap="Switchboard frequency-based information curves"}
switch_results %>% 
  filter(context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

What about context? So far we’ve only discussed the unigram (“contextless”) surprisal metric, considering words on their own without any explicit incorporation of prior context. As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process. We now include one word of context (bigrams) and then two words of context (trigrams). We observe a flattening effect of context across both modalities and all speaker populations. After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.

```{r, echo=F,fig.height=2,fig.cap="Switchboard bigram context-based information curves"}
switch_results %>% 
  filter(context == 1) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")
```

```{r, echo=F,fig.height=2,fig.cap="Switchboard trigram context-based information curves"}
switch_results %>% 
  filter(context == 2) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

We observe the same distinctive distribution of information for parents and children in the Providence corpus as we did for adults in the BNC and Switchboard. The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances.

```{r, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves"}
childes_results %>% 
  filter(language == "English", context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

What about context in the Providence corpus? When incorporating one or two words of predictive context, we observe the same trend as in the Switchboard and BNC corpora. Beyond the first couple of words, once your interlocutor has enough context to predict with some accuracy what you will say next, then you decode information from their speech stream at a constant and optimal rate. This applies to parents and children speaking to one another, as well as adults speaking and writing to one another.

```{r, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
childes_results %>% 
  filter(language == "English", context == 1) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

```{r, echo=F, fig.height =2, fig.cap="Providence trigram context-based information curves"}
childes_results %>% 
  filter(language == "English", context == 2) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package from the R stats core library [@team2013r]. We used the complete linkage algorithm for hierarchical clustering with distances between information curves in language computed using cosine distance between their embeddings in the slope space. The complete linkage algorithm at every step pairs each language or cluster of languages with its closest neighboring language or cluster. A sample from the dendrogram is shown below, and the full tree can be constructed using the code in our GitHub repository. The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages. This suggests using a first-pass quantitative method that the information curves do correspond in some measure to language families, but language families do not explain the entire story.

```{r, echo=F}
wiki_curves %>%
  filter(gram == "Unigram") %>%
  select(-gram) %>% 
  sample_n(40) %>%
  column_to_rownames("language") %>%
  dist() %>% 
  hclust() %>%
  ggdendrogram(rotate=T, leaf_labels = F)
```

A sample of the trigram information curves are plotted below, and all trigram information curves for the languages we used follow the same pattern. The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.

```{r, echo=F, fig.cap="Some trigram information curves from the Wikipedia data"}
wiki_curves %>%
  filter(gram == "Trigram") %>% 
  mutate(Slope0 = runif(1, 0.9, 1.1)) %>% 
  mutate(Slope1 = Slope1 + Slope0) %>% 
  mutate(Slope2 = Slope2 + Slope1) %>% 
  mutate(Slope3 = Slope3 + Slope2) %>% 
  mutate(Slope4 = Slope4 + Slope3) %>% 
  mutate(Slope5 = Slope5 + Slope4) %>% 
  select(-gram) %>% 
  sample_n(6) %>% 
  gather(position, information, -language) %>%
  ggplot(aes(x = position, y = information, group = language, color = language)) + 
    geom_point() + 
    geom_line() + 
    xlab("Word position") + 
    ylab("Average Information") + 
    theme(axis.text.x = element_blank())
```

First we examined the effects of individual typological features on the shapes of the unigram information curves. We ran logistic regressions using the lme4 package in R [@bates2014fitting], checking whether the cosine distance between two languages’ embeddings in the slope space played a role in determining if they had the same value for a given WALS feature. Individual WALS features do not necessarily have ordinal values. Some, such as the “Number of Cases” feature, are easy to quantify and order. Others are more difficult. For example, how does one order “relative clauses appear after the nouns they modify”, “relative clauses appear before the nouns they modify” and “free order of relative clauses and nouns”? We chose the identify relation to avoid deciding on the basis of individual features. We found that 100 out of the 120 features from WALS we examined was statistically significant (p < .001) in determining whether two languages had the same shape to their unigram information distributions. The results for some important features are below, and the rest of the results are in the Appendices.

```{r, echo=F, fig.cap="Some linear model results from Wikipedia and WALS features"}
sf <- wiki_lr %>% 
  filter(p.value == 0, estimate > 0) %>% 
  select(name) %>% 
  rename("Some Important Features" = name) %>% 
  slice(c(1, 10, 12, 15, 23, 29))

@@

We next compared how the cosine distance between two languages related to how many WALS features they had in common. There is a small correlation with a very small R^2 value between WALS features similarity and cosine similarity between two languages, but a statistically significant slope (p < .001). Essentially, for every feature two languages have in common, they have one degree more cosine similarity in the slope space. As feature values tend to be highly correlated within categories (e.g. the word order features are highly correlated), then this translates into language having much more similar information curves if they have more WALS features in common, an intuitive result.

```{r, echo = F, fig.cap="wals features vs cosine similarity"}
wiki_results %>% 
  ggplot(aes(x = feature, y=unigram_cosine)) + 
    geom_smooth(aes(weight = .01), se=F, method = "lm") + 
    geom_hex()
```

So the more similar two languages are in terms of typological features, the more similar their unigram information curves tend to be. For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves.

```{r, echo = F, fig.cap="ldn features vs cosine similarity"}
wiki_results %>% 
  ggplot(aes(x = ldn, y=unigram_cosine)) + 
    geom_smooth(aes(weight = .01), se=F, method = "lm") + 
    geom_hex()
```

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(feather)
library(ggdendro)
library(gridExtra)
library(ggpmisc)

@@

See in the FIGURE below a qualitative comparison of the information trajectory they found and the information trajectory we found. It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as @yu2016distribution did. The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.

@@

For Switchboard we find the same information trajectory as we did for the BNC, and we find the same distribution of information regardless of utterance length. The longer the utterance, the more pronounced the information distribution shape becomes. A qualitative comparison of the two trajectories is in Figure \@ref(fig:switchboard_unigrams).

```{r switchboard_unigrams, echo=F, fig.height=2, fig.cap="Switchboard frequency-based information curves"}
switch_results %>% 
  filter(context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

What about context? So far we’ve only discussed the unigram (“contextless”) surprisal metric, considering words on their own without any explicit incorporation of prior context. As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process. We now include one word of context (bigrams) and then two words of context (trigrams). We observe a flattening effect of context across both modalities and all speaker populations. After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate. The bigram and trigram information curves for Switchboard are in Figures \@ref(fig:switchboard_bigrams) and \@ref(fig:switchboard_trigrams).

```{r switchboard_bigrams, echo=F,fig.height=2,fig.cap="Switchboard bigram context-based information curves"}
switch_results %>% 
  filter(context == 1) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")
```

```{r switchboard_trigrams, echo=F,fig.height=2,fig.cap="Switchboard trigram context-based information curves"}
switch_results %>% 
  filter(context == 2) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

We observe the same distinctive distribution of information for parents and children in the Providence corpus as we did for adults in the BNC and Switchboard. The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances. See \@ref(fig:providence_unigrams).

```{r providence_unigrams, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves"}
childes_results %>% 
  filter(language == "English", context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

What about context in the Providence corpus? When incorporating one or two words of predictive context, we observe the same trend as in the Switchboard and BNC corpora. Beyond the first couple of words, once your interlocutor has enough context to predict with some accuracy what you will say next, then you decode information from their speech stream at a constant and optimal rate. This applies to parents and children speaking to one another, as well as adults speaking and writing to one another. See \@ref(fig:providence_bigrams) and \@ref(fig:providence_trigrams).

```{r providence_bigrams, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
childes_results %>% 
  filter(language == "English", context == 1) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

```{r providence_trigrams, echo=F, fig.height =2, fig.cap="Providence trigram context-based information curves"}
childes_results %>% 
  filter(language == "English", context == 2) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package from the R stats core library [@team2013r]. We used the complete linkage algorithm for hierarchical clustering with distances between information curves in language computed using cosine distance between their embeddings in the slope space. The complete linkage algorithm at every step pairs each language or cluster of languages with its closest neighboring language or cluster. A sample from the dendrogram is shown in Figure \@ref(fig:dendro), and the full tree can be constructed using the code in our GitHub repository. The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages. This suggests using a first-pass quantitative method that the information curves do correspond in some measure to language families, but language families do not explain the entire story.

```{r dendro, echo=F}
wiki_curves %>%
  filter(gram == "Unigram") %>%
  select(-gram) %>% 
  sample_n(40) %>%
  column_to_rownames("language") %>%
  dist() %>% 
  hclust() %>%
  ggdendrogram(rotate=T, leaf_labels = F)
```

A sample of the trigram information curves are plotted in Figure \@ref(fig:wiki_trigrams), and all trigram information curves for the languages we used follow the same pattern. The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.

```{r wiki_trigrams, echo=F, fig.cap="Some trigram information curves from the Wikipedia data"}
wiki_curves %>%
  filter(gram == "Trigram") %>% 
  mutate(Slope0 = runif(1, 0.9, 1.1)) %>% 
  mutate(Slope1 = Slope1 + Slope0) %>% 
  mutate(Slope2 = Slope2 + Slope1) %>% 
  mutate(Slope3 = Slope3 + Slope2) %>% 
  mutate(Slope4 = Slope4 + Slope3) %>% 
  mutate(Slope5 = Slope5 + Slope4) %>% 
  select(-gram) %>% 
  sample_n(6) %>% 
  gather(position, information, -language) %>%
  ggplot(aes(x = position, y = information, group = language, color = language)) + 
    geom_point() + 
    geom_line() + 
    xlab("Word position") + 
    ylab("Average Information") + 
    theme(axis.text.x = element_blank())
```

First we examined the effects of individual typological features on the shapes of the unigram information curves. We ran logistic regressions using the lme4 package in R [@bates2014fitting], checking whether the cosine distance between two languages’ embeddings in the slope space played a role in determining if they had the same value for a given WALS feature. Individual WALS features do not necessarily have ordinal values. Some, such as the “Number of Cases” feature, are easy to quantify and order. Others are more difficult. For example, how does one order “relative clauses appear after the nouns they modify”, “relative clauses appear before the nouns they modify” and “free order of relative clauses and nouns”? We chose the identify relation to avoid deciding on the basis of individual features. We found that 100 out of the 120 features from WALS we examined was statistically significant (p < .001) in determining whether two languages had the same shape to their unigram information distributions. The results for some important features are in Figure \@ref(fig:linear_models), and the rest of the results are in the Appendices.

```{r linear_models, echo=F, fig.cap="Some linear model results from Wikipedia and WALS features"}
sf <- wiki_lr %>% 
  filter(p.value == 0, estimate > 0) %>% 
  select(name) %>% 
  rename("Some Important Features" = name) %>% 
  slice(c(1, 10, 12, 15, 23, 29))

@@

We next compared how the cosine distance between two languages related to how many WALS features they had in common. $r^2$ value is $.005734$, which suggests that in aggregate there is not a correlation between how many WALS features languages have in common and the similarity of their frequency-based information curves. Figure \@ref(fig:cosine_wals) displays the results.

```{r cosine_wals, echo = F, fig.cap="wals features vs cosine similarity"}
r2 <- summary(lm(data = wiki_results, formula = unigram_cosine~feature))$r.squared %>%
  round(digits = 3)
wiki_results %>% 
  ggplot(aes(x = feature, y=unigram_cosine)) + 
    geom_smooth(aes(weight = .01), se=F, method = "lm", color = "black") + 
    geom_jitter(alpha = 0.04, color = "blue") + 
    annotate("text", x = 70, y = -0.5, label = paste0("r^2 = ", r2)) + 
    xlab("Number of identical WALS features") + 
    ylab("Information curve similarity")
    #stat_poly_eq(formula = y~x, position = "dodge")
```

So the more similar two languages are in terms of typological features, the more similar their unigram information curves tend to be. For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves. Figure \@ref(fig:cosine_ldn).

```{r cosine_ldn, echo = F, fig.cap="ldn features vs cosine similarity"}
r2 <- summary(lm(data = wiki_results, formula = unigram_cosine~ldn))$r.squared %>%
  round(digits = 4)
wiki_results %>% 
  ggplot(aes(x = ldn, y=unigram_cosine)) + 
    geom_smooth(aes(weight = .01), se=F, method = "lm", color = "black") + 
    geom_jitter(alpha = 0.04, color = "blue") + 
    annotate("text", x = 0.2, y = -0.5, label = paste0("r^2 = ", r2)) + 
    xlab("Normalized lexical distance") + 
    ylab("Information curve similarity")
    #stat_poly_eq(formula = y~x, position = "dodge")
```

### 2019-09-13 Paper/paper.tex
references and r^2s

  \title{Speakers of diverse languages structure their utterances for efficient communication}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Josef Klafka and Daniel Yurovsky}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{2019-09-11}

@@

See in the figure below a qualitative comparison of the information trajectory they found and the information trajectory we found. It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al. (2016) did. The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.

@@

For Switchboard we find the same information trajectory as we did for the BNC, and we find the same distribution of information regardless of utterance length. The longer the utterance, the more pronounced the information distribution shape becomes. A qualitative comparison of the two trajectories is below.

We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length. This distribution indicates that in English, the words we speak or write at the beginnings of utterances has little information, while the words we speak or write at the ends of utterances has a lot of information. The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from (Genzel \& Charniak, 2002).

What about context? So far we've only discussed the unigram (``contextless'') surprisal metric, considering words on their own without any explicit incorporation of prior context. As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process. We now include one word of context (bigrams) and then two words of context (trigrams). We observe a flattening effect of context across both modalities and all speaker populations. After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.

@@

We observe the same distinctive distribution of information for parents and children in the Providence corpus as we did for adults in the BNC and Switchboard. The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}Providence frequency-based information curves}
\end{figure}

What about context in the Providence corpus? When incorporating one or two words of predictive context, we observe the same trend as in the Switchboard and BNC corpora. Beyond the first couple of words, once your interlocutor has enough context to predict with some accuracy what you will say next, then you decode information from their speech stream at a constant and optimal rate. This applies to parents and children speaking to one another, as well as adults speaking and writing to one another.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}Providence bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}Providence trigram context-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}CHILDES frequency-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}CHILDES bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}CHILDES trigram context-based information curves}
\end{figure}

@@

We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package from the R stats core library (Team \& others, 2013). We used the complete linkage algorithm for hierarchical clustering with distances between information curves in language computed using cosine distance between their embeddings in the slope space. The complete linkage algorithm at every step pairs each language or cluster of languages with its closest neighboring language or cluster. A sample from the dendrogram is shown below, and the full tree can be constructed using the code in our GitHub repository. The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages. This suggests using a first-pass quantitative method that the information curves do correspond in some measure to language families, but language families do not explain the entire story.

\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}

A sample of the trigram information curves are plotted below, and all trigram information curves for the languages we used follow the same pattern. The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{\label{fig:unnamed-chunk-8}Some trigram information curves from the Wikipedia data}
\end{figure}

First we examined the effects of individual typological features on the shapes of the unigram information curves. We ran logistic regressions using the lme4 package in R (Bates, Mächler, Bolker, \& Walker, 2014), checking whether the cosine distance between two languages' embeddings in the slope space played a role in determining if they had the same value for a given WALS feature. Individual WALS features do not necessarily have ordinal values. Some, such as the ``Number of Cases'' feature, are easy to quantify and order. Others are more difficult. For example, how does one order ``relative clauses appear after the nouns they modify'', ``relative clauses appear before the nouns they modify'' and ``free order of relative clauses and nouns''? We chose the identify relation to avoid deciding on the basis of individual features. We found that 100 out of the 120 features from WALS we examined was statistically significant (p \textless{} .001) in determining whether two languages had the same shape to their unigram information distributions. The results for some important features are below, and the rest of the results are in the Appendices.

Some lr results go here.

We next compared how the cosine distance between two languages related to how many WALS features they had in common. There is a small correlation with a very small R\^{}2 value between WALS features similarity and cosine similarity between two languages, but a statistically significant slope (p \textless{} .001). Essentially, for every feature two languages have in common, they have one degree more cosine similarity in the slope space. As feature values tend to be highly correlated within categories (e.g.~the word order features are highly correlated), then this translates into language having much more similar information curves if they have more WALS features in common, an intuitive result.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{\label{fig:unnamed-chunk-9}wals features vs cosine similarity}
\end{figure}

So the more similar two languages are in terms of typological features, the more similar their unigram information curves tend to be. For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{\label{fig:unnamed-chunk-10}ldn features vs cosine similarity}
\end{figure}

>>>>

  \title{Speakers of diverse languages structure their utterances for efficient communication}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Josef Klafka and Daniel Yurovsky}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{2019-09-13}

@@

See in the FIGURE below a qualitative comparison of the information trajectory they found and the information trajectory we found. It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al. (2016) did. The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.

@@

For Switchboard we find the same information trajectory as we did for the BNC, and we find the same distribution of information regardless of utterance length. The longer the utterance, the more pronounced the information distribution shape becomes. A qualitative comparison of the two trajectories is in Figure @ref(fig:switchboard\_unigrams).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/switchboard_unigrams-1.pdf}
\caption{(\#fig:switchboard\_unigrams)Switchboard frequency-based information curves}
\end{figure}

We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length. This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information. The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from (Genzel \& Charniak, 2002).

What about context? So far we've only discussed the unigram (``contextless'') surprisal metric, considering words on their own without any explicit incorporation of prior context. As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process. We now include one word of context (bigrams) and then two words of context (trigrams). We observe a flattening effect of context across both modalities and all speaker populations. After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate. The bigram and trigram information curves for Switchboard are in Figures @ref(fig:switchboard\_bigrams) and @ref(fig:switchboard\_trigrams).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/switchboard_bigrams-1.pdf}
\caption{(\#fig:switchboard\_bigrams)Switchboard bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/switchboard_trigrams-1.pdf}
\caption{(\#fig:switchboard\_trigrams)Switchboard trigram context-based information curves}
\end{figure}

@@

We observe the same distinctive distribution of information for parents and children in the Providence corpus as we did for adults in the BNC and Switchboard. The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances. See @ref(fig:providence\_unigrams).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/providence_unigrams-1.pdf}
\caption{(\#fig:providence\_unigrams)Providence frequency-based information curves}
\end{figure}

What about context in the Providence corpus? When incorporating one or two words of predictive context, we observe the same trend as in the Switchboard and BNC corpora. Beyond the first couple of words, once your interlocutor has enough context to predict with some accuracy what you will say next, then you decode information from their speech stream at a constant and optimal rate. This applies to parents and children speaking to one another, as well as adults speaking and writing to one another. See @ref(fig:providence\_bigrams) and @ref(fig:providence\_trigrams).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/providence_bigrams-1.pdf}
\caption{(\#fig:providence\_bigrams)Providence bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/providence_trigrams-1.pdf}
\caption{(\#fig:providence\_trigrams)Providence trigram context-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}CHILDES frequency-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}CHILDES bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}CHILDES trigram context-based information curves}
\end{figure}

@@

We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package from the R stats core library (Team \& others, 2013). We used the complete linkage algorithm for hierarchical clustering with distances between information curves in language computed using cosine distance between their embeddings in the slope space. The complete linkage algorithm at every step pairs each language or cluster of languages with its closest neighboring language or cluster. A sample from the dendrogram is shown in Figure \ref{fig:dendro}, and the full tree can be constructed using the code in our GitHub repository. The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages. This suggests using a first-pass quantitative method that the information curves do correspond in some measure to language families, but language families do not explain the entire story.

\includegraphics{paper_files/figure-latex/dendro-1.pdf}

A sample of the trigram information curves are plotted in Figure @ref(fig:wiki\_trigrams), and all trigram information curves for the languages we used follow the same pattern. The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/wiki_trigrams-1.pdf}
\caption{(\#fig:wiki\_trigrams)Some trigram information curves from the Wikipedia data}
\end{figure}

First we examined the effects of individual typological features on the shapes of the unigram information curves. We ran logistic regressions using the lme4 package in R (Bates, Mächler, Bolker, \& Walker, 2014), checking whether the cosine distance between two languages' embeddings in the slope space played a role in determining if they had the same value for a given WALS feature. Individual WALS features do not necessarily have ordinal values. Some, such as the ``Number of Cases'' feature, are easy to quantify and order. Others are more difficult. For example, how does one order ``relative clauses appear after the nouns they modify'', ``relative clauses appear before the nouns they modify'' and ``free order of relative clauses and nouns''? We chose the identify relation to avoid deciding on the basis of individual features. We found that 100 out of the 120 features from WALS we examined was statistically significant (p \textless{} .001) in determining whether two languages had the same shape to their unigram information distributions. The results for some important features are in Figure @ref(fig:linear\_models), and the rest of the results are in the Appendices.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/linear_models-1.pdf}
\caption{(\#fig:linear\_models)Some linear model results from Wikipedia and WALS features}
\end{figure}

We next compared how the cosine distance between two languages related to how many WALS features they had in common. \(r^2\) value is \(.005734\), which suggests that in aggregate there is not a correlation between how many WALS features languages have in common and the similarity of their frequency-based information curves. Figure @ref(fig:cosine\_wals) displays the results.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/cosine_wals-1.pdf}
\caption{(\#fig:cosine\_wals)wals features vs cosine similarity}
\end{figure}

So the more similar two languages are in terms of typological features, the more similar their unigram information curves tend to be. For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves. Figure @ref(fig:cosine\_ldn).

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/cosine_ldn-1.pdf}
\caption{(\#fig:cosine\_ldn)ldn features vs cosine similarity}
\end{figure}
